Evolution of pre-copulatory and post-copulatory inbreeding behaviour and associated

### 2017-01-01 ms_pre_post.Rmd
Add bird reference for post-cop avoidance in Intro

Despite a paucity of theory and resulting predictions, numerous empirical studies on diverse species have found evidence for female inbreeding avoidance. However, only a few such studies have determined whether inbreeding avoidance is enacted through pre-copulatory or post-copulatory mechanisms. Among these few studies, pre-copulatory inbreeding avoidance has been documented in weevils, birds, and butterflies [e.g., @Kuriwada2011; @Kingma2013; @Fischer2015], and post-copulatory inbreeding avoidance has been documented in crickets [e.g., @Simmons2006; @Bretman2009]. An even smaller number of studies comprise sufficient evidence for mechanisms of both pre-copulatory and post-copulatory inbreeding avoidance in two well-studied species, including Trinidadian guppies [*Poecilia reticulata*; @Gasparini2011; @Daniel2015] and house mice [*Mus domesticus*; @Potts1991; @Firman2015]. A lack of both pre-copulatory and post-copulatory inbreeding has been found in fruit flies [e.g., @Ala-Honkola2011; @Tan2012], and might be reasonably inferred where no evidence of inbreeding avoidance exists [e.g., @Reid2014; @Reid2015]. To our knowledge, only @Liu2014 consider both pre-copulatory and post-copulatory inbreeding avoidance jointly, finding evidence of pre-copulatory, but not post-copulatory, inbreeding avoidance in cabbage beetles (*Colaphellus bowringi*). Taken together, all of these studies demonstrate that diverse combinations of pre-copulatory and post-copulatory inbreeding avoidance occur in nature. However, there is as yet no theory that predicts what combinations should be favoured by selection when polyandry and pre-copulatory and post-copulatory inbreeding avoidance can all potentially evolve. Consequently, there is no theory that allows the diversity of existing empirical results to be rationalised or interpreted, and no clear hypotheses that could be tested with these studies through comparative analyses or that could be applied to future empirical studies of individual systems.

@@

Evolution of reproductive strategies cannot be presumed to occur in isolation in wild populations, meaning that observed patterns of reproductive strategy within and across taxa are likely to be affected by evolutionary feedbacks and (potentially asymmetric) costs. We have addressed key questions regarding the effect of such feedbacks and costs on evolution of pre-copulatory inbreeding avoidance, post-copulatory inbreeding avoidance, and polyandry. A critical analysis of our model results in new predictions concerning evolutionary dynamics given the joint evolution of these traits, which may be used to formulate new empirical hypotheses within and across taxa.

For example, our results highlight a fundamental but theoretically under-developed conclusion concerning the relationship between evolution of female mating frequency and pre-copulatory versus post-copulatory inbreeding avoidance. Post-copulatory inbreeding avoidance is useless for monandrous females, and most effective for highly polyandrous females that can choose amongst multiple mates to ensure that offspring are fertilised favourably. In contrast, pre-copulatory mechanisms are most salient for monandrous females that have no post-copulatory means for avoiding inbreeding. This conclusion is reflected in our simulation results, as selection for post-copulatory inbreeding avoidance was negligible in monandrous populations, resulting in high drift of allele values over generations due to the inability of females to express post-copulatory choice (e.g., Figure 4C,D). The effect of monandry on selection for pre-copulatory inbreeding avoidance was weaker, but pre-copulatory inbreeding avoidance was generally stronger when populations where monandrous (e.g., Figure 3A vs 3B; see also Supporting Information p. S1-1). In addition to polyandry increasing selection for post-copulatory inbreeding avoidance, polyandry might also covary positively with post-copulatory inbreeding avoidance due to the feed back that post-copulatory inbreeding avoidance has on facilitating evolution of polyandry itself, as observed in our model (Figure 1). It would therefore be interesting to test the hypothesis that, across taxa, occurrence of post-copulatory inbreeding avoidance covaries positively, and pre-copulatory inbreeding avoidance covaries negatively, with degree of polyandry. To test this hypothesis, more work is needed to identify the degree to which females of different species engage in polyandry and the degree to which females express both pre-copulatory and post-copulatory inbreeding avoidance.

Another general inference from our model concerns the dynamics of phenotypes with highly overlapping effects with respect to selection, as realised through functionally different traits. The ability of structurally different phenotypes to perform the same adaptive function is an example of 'degeneracy', a term which defines a more general phenomenon whereby unique elements of a system result in identical outputs [@Edelman2001]. Effects of degeneracy can be observed across nearly all biological scales [@Edelman2001], including complex systems affecting organismal development [e.g., @Nowak1997], adaptation [@Whitacre2010; @Whitacre2010a], and cognition [@Price2002; @Park2013], as well as population [@Atamas2009] and ecosystem [e.g., @Levin2008] dynamics. In the context of reproductive strategy, pre-copulatory inbreeding avoidance, typically expressed through mating behaviour, versus post-copulatory inbreeding avoidance, typically expressed through phenotypes affecting fertilisation, are both driven by the shared selective pressure of inbreeding depression in offspring viability. Hence, degeneracy occurred in our model through an overlap in how phenotypes cause adaptive inbreeding avoidance. Such degeneracy might increase biological robustness [@Gardner2006; @Whitacre2010a], potentially facilitating realised inbreeding avoidance across environments in which either pre-copulatory or post-copulatory mechanisms alone would be ineffective (e.g., due to sexual conflict affecting mate choice or injury affecting fertilisation). However, evolution of each of these strategies might decrease selection on the other strategy by modifying its impact on the degree to which a female avoids inbreeding and consequently has offspring that express inbreeding depression [*sensu* evolution of genetic redundancy; see @Nowak1997]. The relevance of degeneracy with respect to such reproductive strategies therefore requires further theoretical development, which could result in new empirical predictions and conceptual synthesis across biological scales.

>>>>

Despite a paucity of theory and resulting predictions, numerous empirical studies on diverse species have found evidence for female inbreeding avoidance. However, only a few such studies have determined whether inbreeding avoidance is enacted through pre-copulatory or post-copulatory mechanisms. Among these few studies, pre-copulatory inbreeding avoidance has been documented in weevils, birds, and butterflies [e.g., @Kuriwada2011; @Kingma2013; @Fischer2015], and post-copulatory inbreeding avoidance has been documented in birds and crickets [e.g., @Pizzari2004; @Simmons2006; @Bretman2009]. An even smaller number of studies comprise sufficient evidence for mechanisms of both pre-copulatory and post-copulatory inbreeding avoidance in two well-studied species, including Trinidadian guppies [*Poecilia reticulata*; @Gasparini2011; @Daniel2015] and house mice [*Mus domesticus*; @Potts1991; @Firman2015]. A lack of both pre-copulatory and post-copulatory inbreeding has been found in fruit flies [e.g., @Ala-Honkola2011; @Tan2012], and might be reasonably inferred where no evidence of inbreeding avoidance exists [e.g., @Reid2014; @Reid2015]. To our knowledge, only @Liu2014 consider both pre-copulatory and post-copulatory inbreeding avoidance jointly, finding evidence of pre-copulatory, but not post-copulatory, inbreeding avoidance in cabbage beetles (*Colaphellus bowringi*). Taken together, all of these studies demonstrate that diverse combinations of pre-copulatory and post-copulatory inbreeding avoidance occur in nature. However, there is as yet no theory that predicts what combinations should be favoured by selection when polyandry and pre-copulatory and post-copulatory inbreeding avoidance can all potentially evolve. Consequently, there is no theory that allows the diversity of existing empirical results to be rationalised or interpreted, and no clear hypotheses that could be tested with these studies through comparative analyses or that could be applied to future empirical studies of individual systems.

@@

Evolution of reproductive strategies cannot be presumed to occur in isolation in wild populations, meaning that observed patterns of reproductive strategy within and across taxa are likely to be affected by evolutionary feedbacks and (potentially asymmetric) costs. We have addressed key questions regarding the effect of such feedbacks and costs on evolution of pre-copulatory inbreeding avoidance, post-copulatory inbreeding avoidance, and polyandry. A critical analysis of our model engenders new predictions concerning evolutionary dynamics given the joint evolution of these traits, which may be used to formulate new empirical hypotheses within and across taxa.

For example, our results highlight a fundamental but theoretically under-developed relationship between evolution of female mating frequency and pre-copulatory versus post-copulatory inbreeding avoidance. Post-copulatory inbreeding avoidance is useless for monandrous females, and most effective for highly polyandrous females that can choose amongst multiple mates to ensure that offspring are fertilised favourably. In contrast, pre-copulatory mechanisms are most salient for monandrous females that have no post-copulatory means for avoiding inbreeding. This theory is borne out in our simulation results, as selection for post-copulatory inbreeding avoidance was negligible in monandrous populations, resulting in high drift of allele values over generations due to the inability of females to express post-copulatory choice (e.g., Figure 4C,D). The effect of monandry on selection for pre-copulatory inbreeding avoidance was weaker, but pre-copulatory inbreeding avoidance was generally stronger when populations where monandrous (e.g., Figure 3A vs 3B; see also Supporting Information p. S1-1). In addition to polyandry increasing selection for post-copulatory inbreeding avoidance, polyandry might also covary positively with post-copulatory inbreeding avoidance due to the feed back that post-copulatory inbreeding avoidance has on facilitating evolution of polyandry itself, as observed in our model (Figure 1). It would therefore be interesting to test the hypothesis that, across taxa, occurrence of post-copulatory inbreeding avoidance covaries positively, and pre-copulatory inbreeding avoidance covaries negatively, with degree of polyandry. To test this hypothesis, more work is needed to identify the degree to which females of different species engage in polyandry and the degree to which females express both pre-copulatory and post-copulatory inbreeding avoidance.

Another general inference from our simulation results concerns the dynamics of phenotypes with highly overlapping effects with respect to selection, as realised through functionally different traits. The ability of structurally different phenotypes to perform the same adaptive function is an example of 'degeneracy', a term which defines a more general phenomenon whereby unique elements of a system result in identical outputs [@Edelman2001]. Effects of degeneracy can be observed across nearly all biological scales [@Edelman2001], including complex systems affecting organismal development [e.g., @Nowak1997], adaptation [@Whitacre2010; @Whitacre2010a], and cognition [@Price2002; @Park2013], as well as population [@Atamas2009] and ecosystem [e.g., @Levin2008] dynamics. In the context of reproductive strategy, pre-copulatory inbreeding avoidance, typically expressed through mating behaviour, versus post-copulatory inbreeding avoidance, typically expressed through phenotypes affecting fertilisation, are both driven by the shared selective pressure of inbreeding depression in offspring viability. Hence, degeneracy occurred in our model through an overlap in how phenotypes cause adaptive inbreeding avoidance. Such degeneracy might increase biological robustness [@Gardner2006; @Whitacre2010a], potentially facilitating realised inbreeding avoidance across environments in which either pre-copulatory or post-copulatory mechanisms alone would be ineffective (e.g., due to sexual conflict affecting mate choice or injury affecting fertilisation). However, evolution of each of these strategies might decrease selection on the other strategy by modifying its impact on the degree to which a female avoids inbreeding and consequently has offspring that express inbreeding depression [*sensu* evolution of genetic redundancy; see @Nowak1997]. The relevance of degeneracy with respect to such reproductive strategies therefore requires further theoretical development, which could result in new empirical predictions and conceptual synthesis across biological scales.
